The Bot on Speaking Terms: The Effects of Conversation Architecture on Perceptions of Conversational Agents

### 2023-02-15 paper.tex
AK edit of title & abstract

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
% \title[short title]{How to Make Pinocchio a Real Boy? Designing conversational architecture elements to achieve desirable anthropomorphized perceptions}
\title[The Bot with the Eloquent Speech]{The Bot with the Eloquent Speech: Exploring the Relationship between Perceptions of Conversational Agents and Their Conversation Architecture}

@@

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Wei, Kim, and Kuzminykh}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Previous research has demonstrated a relationship between the structure of conversational agents' natural language interaction and how these agents are perceived by their users. However, the systematic connections between the perception of agents and elements of conversation architecture are not well understood. Furthermore, there is no clear definition of the aspects of perception that are affected and the elements of conversation architecture. In this paper, we present a synthesized framework demonstrating the connections between perceptions of agents and conversation architecture elements, using a taxonomy of eleven perceptions grouped into four categories, and ten elements grouped into three categories. We highlight the need for further research for the perception of agent's ability and perception of social connection with agent's, especially as it relates to the linguistic style of conversational agents.  Also, there is the need to be consistent in the assessment of agents' perceptions, and consider ethical concerns in design.
\end{abstract}

@@

Conversational agents (CAs) are designed to interact with humans using natural language through voice or text modalities. Adoption for CAs such as Alexa (Amazon) or Siri (Apple) are rapidly growing in the market, with half of Internet users in United States own one or more smart speaker devices \cite{2022comscore}. These conversational agents are designed to be ubiquitous, easily accessible through smartphones, tablets, laptops and smart devices. Since these agents interact with users using natural language, there are nuances in their communications affected by conversation architecture, which is what they say and how they say it.
%they solicit social responses from users, encouraging users to attribute lifelike qualities, a.k.a. anthropomorphized perceptions to CAs \cite{eyssel2012if}.

%Anthropomorphism is defined as the attribution of traits that we typically associate with being distinctly human to nonhuman entities \cite{waytz2010sees}. For example, users tend to personify conversational agents like Amazon Alexa by using person pronouns instead of object pronouns in online reviews \cite{purington2017alexa}.  Research shows that anthropomorphic design is beneficial to establishing and maintaining trust between users and conversational agents \cite{seeger2021chatbots}. Also, speakers tend to align their lexical choices with conversational agents, similar to human-human conversations \cite{cowan2015does}. Anthropomorphism has been shown to ease user interactions, e.g. making agents more approachable, engaging, and trustworthy when they exhibit human characteristics \cite{qiu2009evaluating}. These processes are of particular interest for human-agent communication because they form the basis for user expectations and predispositions regarding agents' behavior, reliability of the information, etc. \cite{kuzminykh2020genie}.

Previous research has show that there is a connection between conversation architecture and the perception of agents. Studies have found that conversational agents are perceived as more socially present and emotionally intelligent if they use affect language such as sentiment-adaptive response based on user's utterances, resulting in higher user satisfaction \cite{diederich2019emulating}\cite{yang2017perceived}. Also, CAs that mimic user's utterances through lexical alignment reduce users' cognitive workload, resulting in higher engagement with the interaction \cite{spillner2021talk}. Non-linguistic cues such as expressive prosody contribute to higher perceived intimacy with the user, as well as higher perceived enjoyment and ease of use \cite{kim2020can}. Although research on how conversation architecture affects agents' perceptions already exists, there is a lack of understanding on the coverage of relationships explored in literature, as well as synthesizing these findings into a generalized framework. One of the possible issues is the lack of clear and consistent definition on the aspects of user's perceptions of agent and conversation architectural elements.
%However, conversational architecture design may lead to negative consequences, such as an agent with an extroverted personality may be seen as less trustworthy, as users are irritated by the system's overly familiar attitude \cite{andrews2012system}.

@@

For conversation architecture elements, Feine et al \cite{feine2019taxonomy} created a foundational knowledge by providing a taxonomy of social cues, which is defined as "a cue of a CA that triggers a social reaction of the user towards the CA." Elements of conversation architecture are captured in the taxonomy, alongside other social cues such as the embodied agents' appearance and movements. Clark et al \cite{clark2019state} had also captured conversation architecture elements under the research topic of system speech production, with similar categorizations across content, style, and other topics. While these papers captured conversation architecture elements as part of their broader scope of research, there exist the need for standardized definitions specifically for conversation architecture that reflects its unique characteristics.

@@

As shown in Table \ref{tab:cues}, there are three categories defined for the conversation architecture elements covering 10 different themes. The details of each category and theme are discussed below.

>>>>

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
% \title[short title]{How to Make Pinocchio a Real Boy? Designing conversational architecture elements to achieve desirable anthropomorphized perceptions}
\title{The Bot on Speaking Terms: The Effects of Conversation Architecture on Perceptions of Conversational Agents}

@@

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Anon}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
%Conversational agents employ natural language to interact with human users. While it is broadly accepted that the specifics of speech interactions affect the perception of communicating parties, 
%Previous research has demonstrated a relationship between the structure of conversational agents' natural language interaction and how these agents are perceived by their users. However, 
%there is a lack of systematic understanding of the connections between the perception of agents and elements of conversation architecture. % are not well understood. 
%Furthermore, there is no clear definition of the aspects of perception that are affected and the elements of conversation architecture. In this paper, we present a synthesized framework demonstrating the connections between perceptions of agents and conversation architecture elements, using a taxonomy of eleven perceptions grouped into four categories, and ten elements grouped into three categories. We highlight the need for further research for the perception of agent's ability and perception of social connection with agent's, especially as it relates to the linguistic style of conversational agents.  Also, there is the need to be consistent in the assessment of agents' perceptions, and consider ethical concerns in design.
Conversational agents are deliberately designed to interact by simulating natural conversation, and the effectiveness of these interactions is shown to be affected by users' perceptions of an agent. While research on human-human and human-agent communication suggests the association between speech specifics and perceptions of communicating parties, there is a lack of systematic understanding of how an agent’s conversation architecture affects users' perceptions. To address this gap, we present a framework that outlines the relationships between elements of agents’ conversation architecture grouped around dialog strategy, content affectiveness, content style and speech format, and agent characteristics grouped around perceptions of interactions with an agent, agent's ability, sociability, and humanness. Synthesized based on literature review (n=57), the framework demonstrates both the identified relationships and the areas lacking empirical evidence. We discuss the implications of the framework for the conversation design and highlight the inconsistency issues with the domain terminology and measurements.
\end{abstract}

@@

Conversational agents (CAs), such as Amazon Alexa or Siri, are designed to interact with humans using natural language through voice or text modalities. %Adoption of CAs such as Amazon Alexa or Siri (Apple) is rapidly growing in the market, with half of Internet users in United States owning one or more smart speaker devices \cite{2022comscore}. These conversational agents are designed to be ubiquitous, easily accessible through smartphones, tablets, laptops and smart devices. Since these agents interact with users using natural language, there are nuances in their communications affected by conversation architecture, which is what they say and how they say it.
%they solicit social responses from users, encouraging users to attribute lifelike qualities, a.k.a. anthropomorphized perceptions to CAs \cite{eyssel2012if}.
%Anthropomorphism is defined as the attribution of traits that we typically associate with being distinctly human to nonhuman entities \cite{waytz2010sees}. For example, users tend to personify conversational agents like Amazon Alexa by using person pronouns instead of object pronouns in online reviews \cite{purington2017alexa}.  Research shows that anthropomorphic design is beneficial to establishing and maintaining trust between users and conversational agents \cite{seeger2021chatbots}. Also, speakers tend to align their lexical choices with conversational agents, similar to human-human conversations \cite{cowan2015does}. Anthropomorphism has been shown to ease user interactions, e.g. making agents more approachable, engaging, and trustworthy when they exhibit human characteristics \cite{qiu2009evaluating}. These processes are of particular interest for human-agent communication because they form the basis for user expectations and predispositions regarding agents' behavior, reliability of the information, etc. \cite{kuzminykh2020genie}.
Previous research has shown that there is a connection between conversation architecture and the perception of agents. Studies have found that conversational agents are perceived as more socially present and emotionally intelligent if they use affect language such as sentiment-adaptive response based on user's utterances, resulting in higher user satisfaction \cite{diederich2019emulating}\cite{yang2017perceived}. Also, CAs that mimic user's utterances through lexical alignment reduce users' cognitive workload, resulting in higher engagement with the interaction \cite{spillner2021talk}. Non-linguistic cues such as expressive prosody contribute to higher perceived intimacy with the user, as well as higher perceived enjoyment and ease of use \cite{kim2020can}. Although research on how conversation architecture affects agents' perceptions already exists, there is a lack of understanding on the coverage of relationships explored in literature, as well as synthesizing these findings into a generalized framework. One of the possible issues is the lack of clear and consistent definition on the aspects of user's perceptions of agent and conversation architectural elements.
%However, conversational architecture design may lead to negative consequences, such as an agent with an extroverted personality may be seen as less trustworthy, as users are irritated by the system's overly familiar attitude \cite{andrews2012system}.

@@

For conversation architecture elements, Feine et al \cite{feine2019taxonomy} created foundational knowledge by providing a taxonomy of social cues, which is defined as "a cue of a CA that triggers a social reaction of the user towards the CA." Elements of conversation architecture are captured in the taxonomy, alongside other social cues such as the embodied agents' appearance and movements. Clark et al \cite{clark2019state} had also captured conversation architecture elements under the research topic of system speech production, with similar categorizations across content, style, and other topics. While these papers captured conversation architecture elements as part of their broader scope of research, there exist the need for standardized definitions specifically for conversation architecture that reflects its unique characteristics.

@@

As shown in Table \ref{tab:cues}, there are three categories defined for conversation architecture elements across ten different elements. The details of each category and element are discussed below.
